\appchapter[Appendix A]{Speaker Demographics}
\label{ap:A}
\section{Non-native Speaker Demographics}

The 100 audio stimuli were extracted from 93 different non-native English speakers. Information of the non-native speakers' gender, age, age onset of English learning (AO), duration of residence in an English-speaking country (LoR), way of English learning (style) and native language (L1) are available on the website of the Speech Accent Archive. 

Among the 93 non-native English speakers, 41 of them are female and the remaining 52 are male. Age range is between 18 and 69. The mean age is 31.66 (SD=11.39). The AO ranges from 1 to 39. The mean AO is 13.60 (SD=7.87). The LoR ranges from 0 years to 41 years. The mean LoR is 6.28 (SD=8.30). 75 of the 93 non-native speakers learned English in academic settings. 18 of them reported to have acquired English in naturalistic ways.

The non-native speakers differed in their native languages. There are 53 different native languages represented. Table \ref{table:sd1} lists the 53 languages. The numbers in parentheses show the number of speakers of the listed languages.
\figSpace
\begin{longtable}{lllll}
\caption{L1 Background the the L2 English Speakers}\\
\hline
{[}1{]}  & afrikaans (1) & agni (1)      & albanian (2)           & amharic (1)    \\
{[}5{]}  & arabic (4)    & armenian (1)   & baga (1)               & belarusan (1)  \\
{[}9{]}  & bengali (1)   & bosnian (1)    & bulgarian (1)          & cantonese (5)  \\
{[}13{]} & catalan (1)   & czech (3)      & danish (1)             & dutch (1)      \\
{[}17{]} & farsi (2)     & french  (4)    & georgian (2)           & german (4)     \\
{[}21{]} & greek (1)     & gujarati (1)   & italian (3)            & japanese (5)   \\
{[}25{]} & kiswahili (2) & korean (2)     & lamaholot (1)          & mandarin (2)   \\
{[}29{]} & mongolian (1) & norwegian  (3) & polish (1)             & portuguese (4) \\
{[}33{]} & pulaar (1)    & punjabi (2)    & quechua (1)            & romanian (1)   \\
{[}37{]} & russian (2)   & satawalese (1) & serbian (1)            & sinhalese (1)  \\
{[}41{]} & spanish (4)   & swedish (2)    & swiss german (1) &taishan (1)            \\
{[}45{]} & taiwanese (2) & tamil (1)      & telugu (1)             & thai (1)      \\
{[}49{]} & tibetan (1)   & turkish (1)    & uzbek  (1)             & vietnamese (2) \\
{[}53{]} & zulu (1)      &            &                    &       \\
\hline    
\label{table:sd1}
\end{longtable}
\figSpace
\section{Native Speaker Demographics}
\label{dem:native}

Chapter \ref{ch:6} selected 100 native speakers of American English for the construction of the native speaker model. 

Among the 100 native speakers, 39 are female and 61 are male. Age range is between 18 to 79. The mean age is 36.65 (SD=16.93). 97 speaker reported that they reside in the United States. One speaker resides in the United States Virgin Islands. Two speakers reside in both the United States and the United Kingdom.

The 100 native speakers are from 37 states and the District of Columbia in the United States. Table \ref{table:bp3} lists the birthplaces of the native speakers. The numbers in parentheses show the number of speakers from the listed states.
\figSpace
\begin{longtable}{lllll}
\caption{Birthplace of the Native Speakers of American English}\\
\hline
{[}1{]}  & alabama (3)       & alaska (1)         & arkansas (3)             \\
{[}4{]}  & california (9)    & connecticut (2)    & district of columbia (1) \\
{[}7{]}  & florida (4)       & georgia (2)        & idaho (1)                \\
{[}10{]} & illinois (2)      & indiana (3)        & iowa (1)                 \\
{[}13{]} & kansas (1)      & kentucky (1)       & louisiana (3)            \\
{[}16{]} & maine (1)         & maryland (4)      & massachusetts (3)        \\
{[}19{]} & michigan (4)     & minnesota (4)     & mississippi (1)          \\
{[}22{]} & missouri (1)      & new hampshire  (1) & new jersey (1)           \\
{[}25{]} & new york (8)      & north carolina (3) & ohio (7)                 \\
{[}28{]} & pennsylvania (5)  & rhode island (1)   & south carolina (1)       \\
{[}31{]} & south dakota (1)  & texas (4)          & us virgin islands  (1)   \\
{[}34{]} & utah (1)          & virginia (4)      & washington (1)           \\
{[}37{]} & west virginia (1) & wisconsin (3)      &      \\
\hline               
\label{table:bp3}
\end{longtable}



