\appchapter[Appendix B]{Rater Demographics}

\section{Demographics Questionnaire}

Please provide us with some information about you and how you did the experiment. We will keep this information private (it will not be associated with your worker id), and it will help us very much when we analyze the data.
\begin{enumerate}

\item Gender: 

		$\square$ male\quad $\square$ female

\item Age: \underline{\hspace{3cm}}
\item List your native language \underline{\hspace{3cm}}\\
Note: Your native language is the language you heard, grew up with, and used from a young age, typically between the ages 0 and 4. It continues to be the language you identify with, and it is considered your first language. 
\item List any other languages you speak \underline{\hspace{3cm}}
\item Please tell us your occupation \underline{\hspace{3cm}}
\item Please tell us your birth place (city/state/country) \underline{\hspace{3cm}}
\item Please tell us your current residence (city/state/country) \underline{\hspace{3cm}}
\item	Have you ever had any speech or hearing disorders (e.g., Hearing loss, aphasia, dysphasia etc.)? 
		$\square$ yes \quad $\square$ no

\item If Yes, please explain. \\
\underline{\hspace{3cm}}
\item Please provide us with any comments or suggestions you might have regarding this experiment.\\
\underline{\hspace{3cm}}
\end{enumerate}

\section{Experiment 1: Rater Demographics}

Experiment 1 recruited 110 raters. Responses from 108 of the 110 raters were used for the final analysis. As mentioned in Section \ref{rater:1} of Chapter \ref{ch:4}, 61 of the raters are female, 45 are male, and two did not report their gender. The age of the 108 participants ranges from 20 to 66. The mean age is 33.50 (SD=12.51). Table \ref{table:bp1} lists the current residences and birthplaces of these 108 raters. The raters are from 33 states in the United States. The ``Numbers (CR)" column lists the number of raters according to the raters' current residences. The ``Numbers (BP)" column lists the number of raters according to the raters' birthplaces.
 \figSpace
\begin{longtable}{lclc}
  \caption{Raters' Current Residences and Birthplaces (Experiment 1)}\\
\hline
Current Residences & Numbers (CR) & Birthplaces    & Numbers (BP) \\
\hline
arizona           & 4            & arizona        & 4            \\
arkansas          & 1            & arkansas       & 1            \\
california        & 17           & california     & 17           \\
colorado          & 1            & colorado       & 1            \\
connecticut       & 1            & connecticut    & 1            \\
florida           & 8            & florida        & 8            \\
georgia           & 4            & georgia        & 4            \\
illinois          & 1            & illinois       & 1            \\
indiana           & 1            & indiana        & 1            \\
iowa              & 1            & iowa           & 1            \\
kansas            & 2            & kansas         & 2            \\
kentucky          & 2            & kentucky       & 2            \\
louisiana         & 1            & louisiana      & 1            \\
maine             & 1            & maine          & 1            \\
maryland          & 1            & maryland       & 1            \\
massachusetts     & 4            & massachusetts  & 4            \\
michigan          & 2            & michigan       & 2            \\
minnesota         & 1            & minnesota      & 1            \\
missouri          & 3            & missouri       & 3            \\
nevada            & 2            & nevada         & 2            \\
new hampshire     & 1            & new hampshire  & 1            \\
new jersey        & 6            & new jersey     & 6            \\
new mexico        & 1            & new mexico     & 1            \\
new york          & 11           & new york       & 11           \\
north carolina    & 5            & north carolina & 5            \\
ohio              & 3            & ohio           & 3            \\
oklahoma          & 1            & oklahoma       & 1            \\
oregon            & 2            & oregon         & 2            \\
pennsylvania      & 4            & pennsylvania   & 4            \\
south carolina    & 2            & south carolina & 2            \\
tennessee         & 2            & tennessee      & 2            \\
texas             & 7            & texas          & 7            \\
washington        & 5            & washington     & 5           \\
\hline
\label{table:bp1}
\end{longtable}
\figSpace

\section{Experiment 2: Rater Demographics}

155 raters were recruited for Experiment 2. Responses from 133 of the 155 raters were used for final analysis. Only the 133 raters’ information is reported in Table \ref{table:bp2}. Among the 133 raters, 68 are male, 58 are female; seven raters did not report their gender. The mean age of the 133 raters is 38.42 (SD = 11.84). The raters are from 33 states and the District of Columbia in the continental United States. For birthplace and current residence, 12 raters only responded as “usa” without specifiying the city and state.
\figSpace
\begin{longtable}{lclc}
  \caption{Raters' Current Residences and Birthplaces (Experiment 2)}\\
Current Residences   & Numbers (CR) & Birthplace           & Numbers (BP) \\
\hline
alabama              & 2            & alabama              & 1            \\
arizona              & 1            & arizona              & 2            \\
arkansas             & 2            & arkansas             & 0            \\
california           & 15           & california           & 17           \\
colorado             & 1            & colorado             & 3            \\
delaware             & 1            & delaware             & 0            \\
district of columbia & 0            & district of columbia & 2            \\
florida              & 7            & florida              & 2            \\
georgia              & 1            & georgia              & 0            \\
idaho                & 3            & idaho                & 1            \\
illinois             & 4            & illinois             & 4            \\
indiana              & 4            & indiana              & 2            \\
iowa                 & 4            & iowa                 & 2            \\
kansas               & 1            & kansas               & 1            \\
kentucky             & 0            & kentucky             & 1            \\
maryland             & 0            & maryland             & 1            \\
michigan             & 8            & michigan             & 8            \\
minnesota            & 2            & minnesota            & 1            \\
mississippi          & 0            & mississippi          & 1            \\
nevada               & 0            & nevada               & 3            \\
new hampshire        & 0            & new hampshire        & 1            \\
nebraska             & 1            & nebraska             & 0            \\
new jersey           & 8            & new jersey           & 8            \\
new york             & 11           & new york             & 13           \\
north carolina       & 1            & north carolina       & 2            \\
ohio                 & 2            & ohio                 & 5            \\
oklahoma             & 1            & oklahoma             & 1            \\
oregon               & 1            & oregon               & 0            \\
pennsylvania         & 7            & pennsylvania         & 7            \\
rhode island         & 1            & rhode island         & 3            \\
south carolina       & 1            & south carolina       & 3            \\
south dakota         & 0            & south dakota         & 1    \\
tennessee            & 1            & tennessee            & 2            \\
texas                & 8            & texas                & 8            \\
virginia             & 15           & virginia             & 8            \\
washington           & 4            & washington           & 4            \\
west virginia        & 2            & west virginia        & 2            \\
wisconsin            & 1            & wisconsin            & 1            \\
usa                  & 12           & usa                  & 12           \\
\hline       
\label{table:bp2}
\end{longtable}
\figSpace
